\sys: Property-Guided Program Regeneration\\{\vspace{1em}\large Technical University of Denmark, MSc Thesis {\\\vspace{-1em}} Advisors: Christian Gram Kahlauge and Nikos Vasilakis

### 2025-07-22 docs/paper/draft.tex
Small tweaks
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\heading{A compromised javascript library}
The \ttt{flatmap-stream} library was implicated in a high-profile
supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
While the library preserved
its client-observable behavior, an malicious payload inserted in the module's code by an adversarial developer accessed the filesystem, network, and
global state under specific conditions.
Speficially, 
the library would
(1) detect whether it is running under the dependency tree of a popular bitcoin wallet
application~\cite{copay} and on the live bitcoin network,
(2) collect the user's private keys, 
(3) send those to an attacker-controlled machine,
and finally (4) perform its benign functionality.
Program regeneration can automatically remove such malicious payloads
using the insight that the insight that they retain the original library's functionality,
while performing a side-effectful attack on-the-side~\cite{harp:ccs:2021}.
This approach can secure the component even when the attack is highly obfuscated and 
activates under precise conditions.
The slightly simplified source code of the compromised library is as follows:
\begin{minted}[frame=lines]{js}
module.exports = function (stream, fn) {
 if (process.env.PRODUCTION) {
  const priv = copay.getKeys()?.[0]?.priv;
  if (priv && wallet.balance > 0) {
    http.request({ hostname: 'evil.com' })
        .send({ priv });
  }} return flatten(steam).apply(fn); };
\end{minted}

\begin{wrapfigure}[4]{r}{.7\columnwidth}
\vspace{-10pt}
\hspace{-10pt}
\begin{minted}{prolog}
graph(p_, I, O) :- graph(p, I, O).
fsignt(p_, S) :- fsignt(p, S).
pure(p_).
\end{minted}
\end{wrapfigure}
To the right is a program written in the \sys query language to express 
this regeneration goal.
This query requires the regenerated program \ttt{p\_} to exhibit the same I/O behavior as the original program \ttt{p}
but forbids side-effects.
\sys takes this query and arrives at the set of properties to extract from the
original program, given its knowledge base and the query.
This set is $\{\ttt{graph(p, I, O)}, \ttt{inpt(P, I)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}\}$, where
\ttt{input(p, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
by providing the program's source code to an LLM instance and asking it to generate a set of inputs;
\ttt{graph(p, I, O)} extracts the client-observable I/O behavior of the program
by executing the original program on the generated inputs and recording the outputs
(an example I/O pair is $\langle\ttt{([1,[2,3]], (x)=>x+1)}\to\ttt{[2,3,4]}\rangle$);
and \ttt{fsignt(p, S)} extracts the function's signature using a language-aware parser.
The, \ttt{pure(p)} property means that \sys will not provide to the LLM 
instance any information that could encode side-effects, such as the
program's source code, a system call trace, or a filesystem access trace.
After 30.4 seconds and one attempt, having generated 30 I/O pairs, \sys generates the following program:
\begin{minted}[frame=lines]{js}
module.exports = function(stream, fn) {
  return stream.map(fn); };
\end{minted}

@@

\heading{Compilation to ASP}
Given a source program written in the \sys query language, the compiler performs a syntactic
de-sugaring into ASP rules~(\cref{fig:asp-compilation}).

@@

These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.

@@

In the logic space, properties are encoded as predicates over abstract program
identifiers.
To bridge this abstraction gap, \sys given the set of true predicates
from the logic space, extracts the corresponding concrete properties from the
source program.
For example, the predicate $\ttt{inputs}(p, I)$ in the logic space
will be concertized to set of inputs $I$ by invoking the \ttt{input} plugin,
which uses an LLM to generate a set of inputs for the program $p$.
Similarly, the predicate $\ttt{graph}(p, I, O)$ will be concertized to a set of
input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
recording the outputs $O$.
After regeneration, the predicates over the concrete regenerated program $p\_$
will be checked regarding their relationship to the same properties over program $p$
in the logic space.
An equality relationship is one where the predicate's grounded version
$X$ is the same between the two programs, modulo the program identifier~(\eg, $\ttt{graph}(p, i, o) \equiv \ttt{graph}(p\_, i, o)$).
Equality is defined differently for each property, and is enforced by the corresponding plugin.
Similarly, minimization and maximization relationships are defined as
the concrete property value being less than or greater than the original program's value, respectively.
Again, the less-than and greater-than operators are defined by the plugin that verifies the property.

@@

The syntax is similar to \textit{Prolog} and \textit{Datalog}, with code and data being represented
through logical terms.
Collections of logical terms, as seen in
\cref{lst:terms}, can represent the program and its properties.
Rules, expressed using
the \ttt{:-} operator, enable complex reasoning by defining relationships
between atoms.
Choice rules, exemplified in \cref{lst:choice}, allow the ASP
solver to make selections among atoms.
Integrity constraints, as shown in
\cref{lst:constraint}, can restrict invalid answer sets. Optimization
directives, such as \ttt{\#minimize} and \ttt{\#maximize} in
\cref{lst:optimization}, guide the solver to output optimal answer sets,
with respect to a given metric (\eg the number of properties to be extracted from the input program).

>>>>

A user applying \sys would issue the following commands that 
take as input a query (\ttt{-q}), an input file parametrized parametrized over its line numbers (\ttt{-i}), and an output file (\ttt{-o}).
\begin{minted}{shell}
$ kvasir -q qp.pl -i flatmap.js -o flatmap.hs
$ kvasir -q qi.pl -i physics.c:10-27 -o physics.c
$ kvasir -q qm.pl -i app.js:1-30 -o app.js
\end{minted}

\heading{A compromised javascript library}
The \ttt{flatmap-stream} library was implicated in a high-profile
supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
While the library preserved
its client-observable behavior, an malicious payload inserted in the module's code by an adversarial developer accessed the filesystem, network, and
global state under specific conditions.
Speficially, 
the library would
(1) detect whether it is running under the dependency tree of a popular bitcoin wallet
application~\cite{copay} and on the live bitcoin network,
(2) collect the user's private keys, 
(3) send those to an attacker-controlled machine,
and finally (4) perform its benign functionality.
Program regeneration can automatically remove such malicious payloads
using the insight that the insight that they retain the original library's functionality,
while performing a side-effectful attack on-the-side~\cite{harp:ccs:2021}.
This approach can secure the component even when the attack is highly obfuscated and 
activates under precise conditions.
The slightly simplified source code of the compromised library showing the malicious payload and a final
line that performs the benign flatmap functionality is shown below:
\begin{minted}[frame=lines]{js}
module.exports = function (stream, fn) {
 if (process.env.PRODUCTION) {
  const priv = copay.getKeys()?.[0]?.priv;
  if (priv && wallet.balance > 0) {
    http.request({ hostname: 'evil.com' })
        .send({ priv });
  }} return flatten(steam).apply(fn); };
\end{minted}

\begin{wrapfigure}[4]{r}{.7\columnwidth}
\vspace{-10pt}
\hspace{-10pt}
\begin{minted}{prolog}
graph(p_, I, O) :- graph(p, I, O).
fsignt(p_, S) :- fsignt(p, S).
pure(p_).
\end{minted}
\end{wrapfigure}
To the right is a program written in the \sys query language to express 
this regeneration goal.
This query requires the regenerated program \ttt{p\_} to exhibit the same I/O behavior as the original program \ttt{p}
but forbids side-effects.
\sys takes this query and arrives at the set of properties to extract from the
original program, given its knowledge base and the query.
This set is $\{$\ttt{graph(p, I, O)}, \ttt{inpt(P, I)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}$\}$, where
\ttt{input(p, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
by providing the program's source code to an LLM instance and asking it to generate a set of inputs;
\ttt{graph(p, I, O)} extracts the client-observable I/O behavior of the program
by executing the original program on the generated inputs and recording the outputs
(an example I/O pair is $\langle\ttt{([1,[2,3]], (x)=>x+1)}\to\ttt{[2,3,4]}\rangle$);
and \ttt{fsignt(p, S)} extracts the function's signature using a language-aware parser.
The, \ttt{pure(p)} property means that \sys will not provide to the LLM 
instance any information that could encode side-effects, such as the
program's source code, a system call trace, or a filesystem access trace.
After 30.4 seconds and one attempt, having generated 30 I/O pairs, \sys generates the following program:
\begin{minted}[frame=lines]{js}
module.exports = function(stream, fn) {
  return stream.map(fn); };
\end{minted}

@@

\heading{Compilation to ASP}
Given a source program written in the \sys query language, the compiler performs a syntactic
de-sugaring into ASP rules~(\cref{fig:asp-compilation}).
These transformations yield a groundable ASP program that can be evaluated by the backend ASP solver.

@@

@@

In the logic space, properties are encoded as predicates over abstract program
identifiers.
To bridge this abstraction gap, \sys given the set of true predicates
from the logic space, extracts the corresponding concrete properties from the
source program.
For example, the predicate $\ttt{inputs}(p, I)$ in the logic space
will be concertized to set of inputs $I$ by invoking the \ttt{input} plugin,
which uses an LLM to generate a set of inputs for the program $p$.
Similarly, the predicate $\ttt{graph}(p, I, O)$ will be concertized to a set of
input-output pairs $(I, O)$ by executing the program $p$ on the inputs $I$ and
recording the outputs $O$.
After regeneration, the predicates over the concrete regenerated program $p\_$
will be checked regarding their relationship to the same properties over program $p$
in the logic space.
An equality relationship is one where the predicate's grounded version
$X$ is the same between the two programs, modulo the program identifier~(\eg, $\ttt{graph}(p, i, o) \equiv \ttt{graph}(p\_, i, o)$).
Equality is defined differently for each property, and is enforced by the corresponding plugin.
Similarly, minimization and maximization relationships are defined as
the concrete property value being less-than or greater-than the original program's value, respectively.
Again, the less-than and greater-than operators are defined by the plugin that verifies the property.

@@

The syntax is similar to \textit{Prolog} and \textit{Datalog}, with code and data being represented
through logical terms.
Collections of logical terms, as seen in
\cref{lst:terms}, can represent the program and its properties.
Rules, expressed using
the \ttt{:-} operator, define relationships
between atoms.
Choice rules, exemplified in \cref{lst:choice}, allow the ASP
solver to make selections among atoms.
Integrity constraints, as shown in
\cref{lst:constraint}, can restrict invalid answer sets. Optimization
directives, such as \ttt{\#minimize} and \ttt{\#maximize} in
\cref{lst:optimization}, guide the solver to output optimal answer sets,
with respect to a given metric (\eg the number of properties to be extracted from the input program).
